Comparative Evaluation of Spark and Flink Stream Processing

### 2017-02-20 paper/paper.tex
added spark & flink impl sections

\documentclass[]{article}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
%opening

@@

\par Big Data is now one of the hot topics in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing exponentially; generated from social media platforms, mobile phones, IoT and finical transactions. Moreover, recent years showed a rapid development of Big Data frameworks such as Spark and Flink that provide more and efficient data processing capabilities.
 \par The Big Data is a method to process not just only just large datasets (volume), but also that arrives in high rate (velocity), and it comes in all kinds of data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety))\cite{svs}.

@@

\subsection{Data Stream Setup}
Our streaming workloads read input data stream from Kafka. In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircrafts position over time, comprise 22 fields of data such as AircraftID, Date message generated, longitude, latitude, and altitude.
\par In our experiment a data stream producer component reads the ADS-B datasets, then publish it to Kafka cluster to be consumed by the workloads in Spark and Flink. Figure 1 illustrates the data producer and the Kafka cluster setup.
The data stream is portioned into four portions over two server, while the data producer publish the stream records randomly to Kafka partitions.

@@

\begin{itemize}
\item Handling parallel input streams.
\item How to aggregate the state of input stream records.
\item Mange the order of stream records.
\item How to provide and update global data model in stream processing task.
\item Evaluate the performance by measuring  the latency and throughput.
\end{itemize}
We give  in Section 3.3 how Spark and Flink are handling these key aspects and related issues, then we discuss the performance evaluation in Section 4.

\subsubsection{Statistics Computation per Trajectory}
In first stream processing workload, we perform statistics calculation for each trajectory (i.e.,  messages of the same ID) in ADS-B messages stream. As example of computed statistics indexes speed mean, mean of location coordinates, min and max altitude, etc. In this workload, we try to cover 
the parallel receiving of input data stream, state-full stream aggregation, and preserve the order of stream records.

\subsection{Workloads Implementation}
\section{Performance Results}

>>>>

\documentclass[]{article}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage[inline]{enumitem}
\usepackage{listings}
%opening

@@

\par Big Data is now one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing exponentially; generated from social media platforms, mobile phones, IoT and finical transactions. Moreover, recent years showed a rapid development of Big Data frameworks such as Spark and Flink that provide more and efficient data processing capabilities.
 \par The Big Data is a method to process not just only just large datasets (volume), but also that arrives in high rate (velocity), and it comes in all kinds of data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety))\cite{svs}.

@@

\subsection{Data Stream Setup}
Our streaming workloads read input data stream from Kafka. In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircrafts position over time, comprise 22 fields of data such as aircraft ID, date message generated, longitude, latitude, and altitude.
\par In our experiment a data stream producer component reads the ADS-B datasets, then publish it to Kafka cluster to be consumed by the workloads in Spark and Flink. Figure 1 illustrates the data producer and the Kafka cluster setup.
The data stream is portioned into four portions over two server, while the data producer publish the stream records randomly to Kafka partitions.

@@

\begin{itemize}
\item Handling parallel input streams.
\item How to aggregate the state of input stream records.
\item Mange the order of stream records.
\item How to provide and update global data model in stream processing task.
\item Evaluate the performance by measuring  the latency and throughput.
\end{itemize}
We present how Spark and Flink are handling these key aspects and related issues in each workload, then we discuss the performance evaluation in Section 4.

\subsubsection{Statistics Computation per Trajectory}
In first stream processing workload, we perform statistics calculation for each trajectory (i.e.,  messages of the same ID) in ADS-B messages stream. As example of computed statistics indexes speed mean, mean of location coordinates, min and max altitude, etc. In this workload, we try to cover 
the parallel receiving of input data stream, state-full stream aggregation, and preserve the order of stream records.
The implementation details of this workload in Spark and Flink, are presented in the following: 
\begin{itemize}
\item {\bf{Implementation in Spark }}

The statistics calculation workload is implemented in Spark by performing the following steps: \begin{enumerate*}[label={\alph*)}]
\item read parallel Kafka input streams 
\item parse the multiple streams records   
\item filter only the relevant  messages 
\item combine the all messages related to the same trajectory
\item manage the aggregate state and compute the statistics for each trajectory.
\end{enumerate*}

\par The used Spark Streaming API operations to execute those steps are shown in Figure 2. First, in Spark to consume parallel Kafka stream, multiple Kafka stream receivers must be created then combine them using the stream \texttt{union} API. Second, we parse each stream record (i.e., ADS-B message) then map it to tuple of aircraft ID and aircraft position object through \texttt{mapToPair} transformation API. Third, the irrelevant messages are filtered by the \texttt{filter}. Fourth, the tuples with same ID are grouped together using the \texttt{groupByKey} Spark API function, after applying this transformation the input stream records are shuffled which means to preserve the order of records a sort operation is needed. Finally, due to the mini-batch nature of Spark streaming, the input stream is processed batch by batch, we use the  \texttt{updateStateByKey} to compute statistics per trajectory after sorting the list of positions, then update the aggregated state of the trajectories stream.

  \centering

     \caption{Spark streaming implementation of statistics computation workload.}
\end{figure}

\item {\bf{Implementation in Flink }}

The statistics computation of trajectories stream is solved in Flink by performing the following steps: \begin{enumerate*}[label={\alph*)}]
\item read Kafka input stream
\item parse the input stream records   
\item filter only the relevant  messages 
\item combine the all messages related to the same trajectory
\item compute the statistics for each trajectory.
\end{enumerate*}

\par The utilized Flink APIs to perform those steps are shown in Figure 3. First, create Kafka stream consumers that run in multiple parallel instances based on the configured \texttt{parallelism} factor. Second,  stream elements (i.e., ADS-B message) are mapped  to tuples of aircraft ID and  position object through \texttt{map} transformation API. Third, the irrelevant messages are filtered by the \texttt{filter}. Fourth, tuples with same ID are combined together using the \texttt{keyBy} API function. Finally, the \texttt{reduce} API function is used to calculate the statistics and aggregate it.

  \centering

     \caption{Flink implementation of statistics computation workload.}
\end{figure}

//TODO: discuss differences between flink and spark.

\end{itemize}
\subsection{Workloads Implementation}
\section{Performance Results}
